A Typed Programmatic Interface to Contracts on the Blockchain

### 2021-02-16 papers/execution-model/model.tex
sec 2 and 3

@@

This code is idealized in several respects. Originating or running a contract
requires to propose a fee to the blockchain, which may or may not be
accepted.
Starting a contract may also time out for a variety of reasons. So just
invoking a contract with a fixed fee does not guarantee the contract's
execution.  Even if the invocation is locally accepted, it still takes
a couple of cycles before we can be sure the invocation is globally
accepted and incorporated in the blockchain.
Hence, after starting the invocation, we have to observe the fate of
this invocation. If it does not get incorporated, then we need to
analyse the reason and react accordingly. For example, if the
invocation was rejected because of a insufficient fee, we might want
to restart with an increased fee. Or we might decide to wait until the
invocation goes through with the low fee, if it is not that important.

Hence, we would implement a scheme similar to
the bidding strategy: start with a low fee and increase until the
contract is accepted or a fee cap is reached. On the other hand, an
observer function like \lstinline/get_state/ always succeeds.

The low-level interface that we propose in this paper requires 
(From this scenario we see the required functionality for a
programmatic interface to a blockchain:)

\begin{itemize}
\item query certain components like the state of a contract (to obtain
  the current highest bid, to check the fees of the last block)
\item invoke contracts and initiate transfers
\end{itemize}

\section{Definitions}
\begin{definition}%[Set]
Let $\mathbf{Elt}$ be the set of (concrete) elements. Let $\emptyset$ be an empty set and $\mathbf{e}$ $\in$  $\mathbf{Elt}$. A set of elements is expressed as the following syntax:
\\
$\mathbf{s} :: = \emptyset \mid \mathbf{e} \mid \mathbf{s :: s}$
\end{definition}

>>>>

\subsection{Contributions}
\label{sec:contributions}

\begin{itemize}
\item An operational semantics for functional programs running
  alongside smart contracts in a blockchain. 
\item Proofs of various properties of the combined system.
\item An implementation of a low-level OCaml-API to the Tezos
  blockchain, which corresponds to the operational semantics.
\end{itemize}

@@

This code is idealized in several respects. Originating or running a contract
requires to propose a fee to the blockchain, which may or may not be
accepted.
Starting a contract may also time out for a variety of reasons. So just
invoking a contract with a fixed fee does not guarantee the contract's
execution.  Even if the invocation is locally accepted, it still takes
a couple of cycles before we can be sure the invocation is globally
accepted and incorporated in the blockchain.
Hence, after starting the invocation, we have to observe the fate of
this invocation. If it does not get incorporated, then we need to
analyse the reason and react accordingly. For example, if the
invocation was rejected because of an insufficient fee, we might want
to restart with an increased fee. Or we might decide to wait until the
invocation goes through without increasing the fee.

Hence, we would implement a scheme similar to
the bidding strategy: start with a low fee and increase (or wait) until the
contract is accepted or a fee cap is reached. On the other hand, an
observer function like \lstinline/get_state/ always succeeds.

The low-level interface that we propose in this paper requires the
programmer to be explicit about fees, waiting, and polling the state
of contract invocations.

In summary, a useful smart-contract-API  has facilities to
\begin{itemize}
\item query the current state of the blockchain (e.g., fees in the
  current block),
\item query the state and the balance of a contract (e.g., to obtain
  the current highest bid),
\item originate contracts, invoke contracts, and initiate
  transfers. To this end, the program using the API has to run on
  behalf of some authority (i.e., expressed by holding the private key
  for some account).
\end{itemize}
All these facilities are supported by the (untyped) RPC interface of the Tezos
blockchain, which is the basis of our implementation.

\section{Execution Model}
\label{sec:execution-model}

The execution model is an abstraction of the Tezos blockchain \cite{Tezos}
augmented with functional (OCaml) programs that interact with the
blockchain. As the blockchain is realized by a peer-to-peer network of
independent nodes, interaction happens through
\emph{local nodes} that receive requests to originate and invoke
contracts from programs that run on a particular node. We model the blockchain
itself as a separate, abstract global entity that represents the
current consensual state of the system. Our model does not express
low-level details, but relies on nondeterminism to describe
the possible behaviors of the system.

We write $\emptyset$ for the empty set and $\mathbf{e :: s}$ to
decompose a set into an element $\mathbf{e}$ and a set  $\mathbf{s}$.
